A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2022-08-22 Papers/Paper/manuscript.Rmd
Roger took revision pass through ms

classoption: [times, biblatex]
#
# Possible classoptions:
# - [times] for Times font (default if no option is chosen)
# - [cm] for Computer Modern font
# - [lucida] for Lucida font (not freely available)
# - [brill] open type font, freely downloadable for non-commercial use from
#   http://www.brill.com/about/brill-fonts; requires xetex
# - [charis] for CharisSIL font, freely downloadable from http://software.sil.org/charis/
# - for the Brill an CharisSIL fonts, you have to use the XeLatex
#   typesetting engine (not pdfLatex)
# - for headings, tables, captions, etc., Fira Sans is
#   used: https://www.fontsquirrel.com/fonts/fira-sans
# - [biblatex] for using biblatex (the default is natbib, DO NOT LOAD the
#   natbib package in this file, it is loaded automatically via the document
#   class glossa.cls)
# - [linguex] loads the linguex example package
#
# !!! A note on the use of linguex: in glossed examples, the third line of the
# example (the translation) needs to be prefixed with \glt. This is to allow a
# first line with the name of the language and the source of the example. See
# example (2) in the text for an illustration.
#
# !!! A note on the use of bibtex: for PhD dissertations to typeset correctly
# in the references list, the Address field needs to contain the city
# (for US cities in the format "Santa Cruz, CA")
#
bibliography: ["r-references.bib","refs.bib"]
biblio-style: "apalike"
link-citations: true
# The bibliography style is set automatically by glossa.cls when using
# either natbib or biblatex.
# Pandoc's citeproc is not supported.
#
abstract: |
   Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing. A-maze is a recent method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities in individual sentences. We adapted A-maze for use on longer passages and tested it on the Natural Stories corpus. Participants were able to comprehend these longer text passages that they read via the Maze task.  Moreover, the Maze task yielded useable reaction time data with word predictability effects that were linearly related to surprisal, the same pattern found with other incremental methods. Crucially, Maze reaction times show a tight relationship with properties of the current word, but little spillover of effects from previous words. This superior localization is an advantage of Maze compared with other methods. Overall, we expanded the scope of experimental materials, and thus theoretical questions, that can be studied with the Maze task. 
# Specify keywords here:
keywords: stylesheet, glossa, article
# To add the word count, uncomment the following option and replace the number
# with the document word count.
# wordcount: 1000
output: 
  bookdown::pdf_book:
    base_format: rticles::glossa_article
    number_sections: yes
---

@@

A major feature of human language is that not all sentences or utterances are equally easy to successfully comprehend. Sometimes this is mostly or entirely due to the linguistic structure of the sentence: for example, *The rat that the cat that the dog chased killed ate the cheese* is more difficult than *The rat that was killed by the cat that was chased by the dog ate the cheese* even though the meaning of the two sentences is (near-)identical. Sometimes the source of difficulty can be a mismatch between expectations set up by the context and the word choice in an utterance: for example, the question *Is the cup red?* may be confusing in a context containing more than one cup. Psycholinguistic theories may differ in their ability to predict what is easy and what is hard. One of the most powerful methods for studying these differential difficulty effects is to turn over control of presentation of the linguistic material to comprehender, and to measure what she takes time on. For this purpose, taking measurements from experimental participants during reading, a widespread, highly practiced skill in diverse populations around the world, is of unparalleled value.

@@

Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually control the visual presentation of the text by pressing a button. In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time, wihth every press of the button masks the currently presented word (group) and simultaneously reveals the next. The time spent between button presses is the unique RT measure for that word (group). Self-paced reading requires no special equipment and can be delivered remotely, but the measurements are noisier and even more prone to spillover [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006; @smithEffectWordPredictability2013].

## Maze
The Maze task is an alternative method that is designed to increase localization at the expense of naturalness [@freedman85; @forsterMazeTaskMeasuring2009]. In the Maze task, participants must repeatedly choose between two simultaneously presented options: a correct word that continues the sentence, and a distractor string which does not. Participants must choose the correct word, and their time to selection is treated as the reaction time, or RT. (We intentionally overload the abbreviation "RT" and use it for Maze reaction times as well as reading times from eye tracking and SPR, because the desirable properties of reading times turn out to hold for Maze reaction times as well.) @forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical "L"-maze where the distractors are non-word strings, and grammatical "G"-maze where the distractors are real words that don't fit with the context of the sentence. In theory, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word. @forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental processing" to distinguish from other incremental processing methods where words can be passively read before later committing to a parse. This idea of strong localization is supported by studies finding strongly localized effects for G-maze [@witzelComparisonsOnlineReading2012a; @boyceMazeMadeEasy2020].

@@

The overall predictability of a word in a context can be estimated using language models that are trained on large corpora of language to predict what word comes next in a sentence. A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability. Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time etc).

@@

However, terminating sentences on errors means that we don't have RTs for words after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, errors are rare and data loss is a small issue. However, this data loss is much worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

@@

## SPR-Maze correlation
We compared the correlations between the Maze and SPR RTs to within-Maze and within-SPR correlations. For Maze, within each story, we randomly split subjects into two halves. Within each half, we calculated a per-word average RT for each word and then a per-sentence average RT across word averages. We calculated a within-Maze correlation between these two halves.

@@

Our analytic questions required multiple modeling approaches. To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships. GAMs are harder to interpret than linear models, so to measure effect sizes and assess spillover, we used linear mixed models. Finally, in order to determine which language model best predicts the RT data, we fit additional linear models with predictors from multiple language models to look at their relative contributions. All these models used surprisal, frequency, and length as predictors for RT. We considered these predictors from both the current and past word to account for the possibility of spillover effects in A-maze. For SPR comparisons, we included predictors from the current and past three words to account for known spillover effects. We conducted data processing and analyses using `r a$r`.

@@

For linear models of the SPR data, we were unable to fit the full mixed effects structure. The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this structure did not fit reliably in brm.

For model comparison, we took by-item averaged data to aid in fast model fitting.  We included frequency, length, and their interaction in all models. Then we fit models with either 1 or 2 sources of surprisal using lm [@batesFittingLinearMixedeffects2015] and assessed the effect of adding the second surprisal source with an anova.

@@

The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. These differences cannot be directly attributed to methods, as the participant populations differed. While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected [@chmielewski2020].

@@

(ref:scatter) Correlation between SPR and Maze data. RTs were averaged across participants per word and then averaged together within each sentence. RTs in ms.

@@

both <- maze %>% 
  group_by(Story_Num, Sentence_Num, word_num) %>% 
  summarize(maze=mean(rt, na.rm=T)) %>% 
  group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=mean(maze)) %>% 
  inner_join(spr_collapse)
ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)+geom_smooth(method="lm")
```

@@

One of the main claimed advantages of the Maze task is that it has better localization and less spillover than SPR. We examined how much spillover A-maze and SPR each had by fitting linear models with predictors from current and previous words. Large effects from previous words are evidence for spillover; effects of the current word that dwarf any lagged effects is evidence for localization.

We modeled reading time as a function of surprisal, frequency, and length as well as surprisal x length and frequency x length interactions. For all of these, we included the predictors for the current and previous word, and we centered, but did not rescale, all predictors. (See Methods for more details on these predictors and model fit process.) As with the GAM models, we used surprisal calculations from 4 different language models for robustness.

@@

Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character. This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms. Thus Maze is is 2-3 times as slow as SPR but has roughly 10 times larger effects.

@@

We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).  For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models.

@@

Fourth, we compared the spillover profiles between Maze and SPR. For Maze, we found large effects of the current words surprisal and length, which dwarfed any spillover effects from previous word predictors. In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words. Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus and thus generating less noisy data.

Lastly, we examined how different language models fare at predicting human RT data. We found that overall, the models were more predictive of the A-maze data than SPR data; however, the hierarchy of the model's predictive performance also differed between the A-maze and SPR datasets. This difference suggests that how predictive a language model is of RTs may depend on the task.  Further comparisons between different processing methods on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

Overall, A-maze has good localization, although some  models showed small but statistically reliable effects of the past word. On the whole, however, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.

@@

Error-correction Maze  reduces perverse incentives from the desire to complete the task quickly compared to traditional Maze, but clicking randomly is still more efficient than doing the task. In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake. This delay would make randomly hitting buttons slower than doing the task as intended.

Error-correction Maze records RTs for words after a participant makes a mistake in the sentence. In our analyses, we excluded this post-error data, but we believe it is an open question whether data from after a participant makes a mistake is usable. That is, does it show the same profile as RTs from pre-error words, or are there traces from recovering from the mistake, and if so, how long do these effects take to fade? Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is hard-to-assess; if it can be used, it would make the Maze task more data efficient.

The Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest. The extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.  We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods, and we suggest that the error-correction mode be the default choice for presenting Maze materials.

>>>>

classoption: [times, biblatex]
#
# Possible classoptions:
# - [times] for Times font (default if no option is chosen)
# - [cm] for Computer Modern font
# - [lucida] for Lucida font (not freely available)
# - [brill] open type font, freely downloadable for non-commercial use from
#   http://www.brill.com/about/brill-fonts; requires xetex
# - [charis] for CharisSIL font, freely downloadable from http://software.sil.org/charis/
# - for the Brill an CharisSIL fonts, you have to use the XeLatex
#   typesetting engine (not pdfLatex)
# - for headings, tables, captions, etc., Fira Sans is
#   used: https://www.fontsquirrel.com/fonts/fira-sans
# - [biblatex] for using biblatex (the default is natbib, DO NOT LOAD the
#   natbib package in this file, it is loaded automatically via the document
#   class glossa.cls)
# - [linguex] loads the linguex example package
#
# !!! A note on the use of linguex: in glossed examples, the third line of the
# example (the translation) needs to be prefixed with \glt. This is to allow a
# first line with the name of the language and the source of the example. See
# example (2) in the text for an illustration.
#
# !!! A note on the use of bibtex: for PhD dissertations to typeset correctly
# in the references list, the Address field needs to contain the city
# (for US cities in the format "Santa Cruz, CA")
#
bibliography: ["r-references.bib","refs.bib"]
biblio-style: "apalike"
link-citations: true
# The bibliography style is set automatically by glossa.cls when using
# either natbib or biblatex.
# Pandoc's citeproc is not supported.
#
abstract: |
   Behavioral measures of word-by-word reading time provide experimental evidence to test theories of language processing. A-maze is a recent method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities in individual sentences. We adapted A-maze for use on longer passages and tested it on the Natural Stories corpus. Participants were able to comprehend these longer text passages that they read via the Maze task.  Moreover, the Maze task yielded useable reaction time data with word predictability effects that were linearly related to surprisal, the same pattern found with other incremental methods. Crucially, Maze reaction times show a tight relationship with properties of the current word, with little spillover of effects from previous words. This superior localization is an advantage of Maze compared with other methods. Overall, we expanded the scope of experimental materials, and thus theoretical questions, that can be studied with the Maze task. 
# Specify keywords here:
keywords: stylesheet, glossa, article
# To add the word count, uncomment the following option and replace the number
# with the document word count.
# wordcount: 1000
output: 
  bookdown::pdf_book:
    base_format: rticles::glossa_article
    number_sections: yes
---

@@

A major feature of human language processing is that not all sentences or utterances are equally easy to successfully comprehend. Sometimes this is mostly or entirely due to the linguistic structure of the sentence: for example, *The rat that the cat that the dog chased killed ate the cheese* is more difficult than *The rat that was killed by the cat that was chased by the dog ate the cheese* even though the meaning of the two sentences is (near-)identical. Sometimes the source of difficulty can be a mismatch between expectations set up by the context and the word choice in an utterance: for example, the question *Is the cup red?* may be confusing in a context containing more than one cup. Psycholinguistic theories may differ in their ability to predict what is easy and what is hard. One of the most powerful methods for studying these differential difficulty effects is to turn over control of presentation of the linguistic material to comprehender, and to measure what she takes time on. For this purpose, taking measurements from experimental participants during reading, a widespread, highly practiced skill in diverse populations around the world, is of unparalleled value.

@@

Self-paced reading (SPR; @mitchell:1984) is a somewhat less natural paradigm in which the participant manually control the visual presentation of the text by pressing a button. In its generally preferred variant, moving-window self-paced reading, words are revealed one at a time or one group at a time: every press of the button masks the currently presented word (group) and simultaneously reveals the next. The time spent between button presses is the unique RT measure for that word (group). Self-paced reading requires no special equipment and can be delivered remotely, but the measurements are noisier and even more prone to spillover [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006; @smithEffectWordPredictability2013].

## Maze
The Maze task is an alternative method that is designed to increase localization at the expense of naturalness [@freedman85; @forsterMazeTaskMeasuring2009]. In the Maze task, participants must repeatedly choose between two simultaneously presented options: a correct word that continues the sentence, and a distractor string which does not. Participants must choose the correct word, and their time to selection is treated as the reaction time, or RT. (We deliberately overload the abbreviation "RT" and use it for Maze reaction times as well as reading times from eye tracking and SPR, because the desirable properties of reading times turn out to hold for Maze reaction times as well.) @forsterMazeTaskMeasuring2009 introduced two versions of the Maze task: lexical "L"-maze where the distractors are non-word strings, and grammatical "G"-maze where the distractors are real words that don't fit with the context of the sentence. In theory, participants must fully integrate each word into the sentence in order to confidently select it, which may require mentally reparsing previous material in order to allow the integration and selection of a disambiguating word. @forsterMazeTaskMeasuring2009 call this need for full integration "forced incremental processing" to distinguish from other incremental processing methods where words can be passively read before later committing to a parse. This idea of strong localization is supported by studies finding strongly localized effects for G-maze [@witzelComparisonsOnlineReading2012a; @boyceMazeMadeEasy2020].

@@

The overall predictability of a word in a context can be estimated using language models that are trained on large corpora of language to predict what word comes next in a sentence. A variety of pre-trained models exist, with varied internal architectures and training methods, but all of them generate measures of predictability. Predictability is often measured in bits of surprisal, which is the negative log probability of a word (1 bit of surprisal means a word is expected to occur half the time, 2 bits is 1/4 of the time, etc.).

@@

However, terminating sentences on errors means that we don't have RTs for words after a participant makes a mistake in an item. In traditional G-maze tasks, with hand-crafted distractors and attentive participants, errors are rare and data loss is a small issue. However, this data loss can be worse with A-maze materials and crowd-sourced participants [@boyceMazeMadeEasy2020]. The high errors are likely from some combination of participants guessing randomly and from auto-generated distractors that in fact fit the sentence; as @boyceMazeMadeEasy2020 noted, some distractors, especially early in the sentence, were problematic and caused considerable data loss.

@@

## SPR--Maze correlation
We compared the correlations between the Maze and SPR RTs to within-Maze and within-SPR correlations. For Maze, within each story, we randomly split subjects into two halves. Within each half, we calculated a per-word average RT for each word and then a per-sentence average RT across word averages. We calculated a within-Maze correlation between these two halves.

@@

Our analytic questions required multiple modeling approaches. To look at the functional form of the relationship between surprisal and RT data, we fit Generalized Additive Models (GAMs) to allow for non-linear relationships. GAM model summaries can be harder to interpret than those for linear models, so to measure effect sizes and assess spillover, we used linear mixed models. Finally, in order to determine which language model best predicts the RT data, we fit additional linear models with predictors from multiple language models to look at their relative contributions. All these models used surprisal, frequency, and length as predictors for RT. We considered these predictors from both the current and past word to account for the possibility of spillover effects in A-maze. For SPR comparisons, we included predictors from the current and past three words to account for known spillover effects. We conducted data processing and analyses using `r a$r`.

@@

For linear models of the SPR data, we were unable to fit a single model whose random effects structure was maximal with respect to all fixed-effects predictors. We report results for the best (in terms of having maximal random effects structure with respect to fixed effects of primary theoretical interest) single model we could fit by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and a per-word-token random intercept, fit with lme4 [@batesFittingLinearMixedeffects2015], as this model specification did not fit reliably in brm.

For model comparisons, we took by-item averaged data to aid in fast model fitting.  We included frequency, length, and their interaction in all models. Then we fit simple linear regression models (using `R`'s `lm()`) with either 1 or 2 sources of surprisal and assessed the effect of adding the second surprisal source with an F test (using `R`'s `anova()`).

@@

The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. These differences cannot necessarily be attributed to methods, as the participant populations differed. While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected [@chmielewski2020].

@@

(ref:scatter) Correlation between SPR and Maze data. RTs (measured in milliseconds) were averaged across participants per word and then averaged together within each sentence, so that each point represents the average RT in the two methods for one sentence in the corpus. Presented on a fixed scale coordinate system where 1 millisecond of RT takes equal physical space on both axes. Line and confidence interval reflect best linear fit regression of SPR time against Maze time.

@@

both <- maze %>% 
  group_by(Story_Num, Sentence_Num, word_num) %>% 
  summarize(maze=mean(rt, na.rm=T)) %>% 
  group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=mean(maze)) %>% 
  inner_join(spr_collapse)
ggplot(both, aes(x=maze, y=spr)) + 
  geom_point(alpha=.5, size=1) + 
  labs(x="A-Maze", y="SPR") + 
  coord_fixed(ratio=1) +
  geom_smooth(method="lm")

# mod1 <- lm(spr ~ maze,both)
# mod2 <- lm(maze ~ spr,both)
# slope2 <- 1/coef(mod2)[2]
# intercept2 <- with(both,mean(spr) - slope2 * mean(maze))
# matr <- with(both,matrix(c(var(maze),cov(spr,maze),cov(spr,maze),var(spr)),2,2))
# slope <- eigen(matr)$vectors[2,1]/eigen(matr)$vectors[1,1]
# intercept <- with(both,mean(spr) - slope * mean(maze))
# ggplot(both, aes(x=maze, y=spr)) + 
#   geom_point(alpha=.5, size=1) + 
#   labs(x="A-Maze", y="SPR") + 
#   #theme(aspect.ratio=1) +
#   coord_fixed(ratio=1) +
#   stat_ellipse(type="norm",segments=501) +
#   stat_ellipse(type="t",color="yellow",segments=501) +
#   #geom_smooth(method="lm") 
#   geom_abline(intercept=coef(mod1)[1],slope=coef(mod1)[2],color="blue") + 
#   geom_abline(intercept=intercept2,slope=slope2,color="red") + 
#   geom_abline(intercept=intercept,slope=slope,color="green")
#   #geom_abline(intercept=0,slope=1)
# 
# slope1 <- 1/coef(mod1)[2]
# intercept1 <- with(both,mean(maze) - slope1 * mean(spr))
# slopeA <- eigen(matr)$vectors[1,1]/eigen(matr)$vectors[2,1]
# interceptA <- with(both,mean(maze) - slopeA * mean(spr))
# ggplot(both, aes(x=spr, y=maze)) + 
#   geom_point(alpha=.5, size=1) + 
#   labs(y="A-Maze", x="SPR") + 
#   theme(aspect.ratio=1) +
#   #coord_fixed(ratio=1) +
#   stat_ellipse(type="norm",segments=501) +
#   stat_ellipse(type="t",color="yellow",segments=501) +
#   #geom_smooth(method="lm") 
#   geom_abline(intercept=coef(mod2)[1],slope=coef(mod2)[2],color="red") + 
#   geom_abline(intercept=intercept1,slope=slope1,color="blue") + 
#   geom_abline(intercept=interceptA,slope=slopeA,color="green")
#   #geom_abline(intercept=0,slope=1)
```

@@

One of the main claimed advantages of the Maze task is that it has better localization and less spillover than SPR. We examined how much spillover A-maze and SPR each had by fitting linear models with predictors from current and previous words. Large effects from previous words are evidence for spillover; effects of the current word dwarfing any lagged effects would be evidence for localization.

We modeled reading time as a function of surprisal, frequency, and length as well as surprisal$\times$length and frequency$\times$length interactions. For all of these, we included the predictors for the current and previous word, and we centered, but did not rescale, all predictors. (See Methods for more details on these predictors and model fit process.) As with the GAM models, we used surprisal calculations from 4 different language models for robustness.

@@

Whereas Maze showed surprisal effects in the 10 to 25 ms/bit range and length effects 15 to 20 ms/character range, SPR effects are about 1-2 ms per bit or character. This difference in effect size is disproportionate to the overall speed of the methods; the predicted intercept for the Maze task was roughly 880 ms and for SPR was roughly 360 ms. Thus Maze is 2--3 times as slow as SPR but has roughly 10 times larger effects.

@@

We followed the same process for the SPR data with results shown in Table \@ref(tab:spr-compare).  For SPR, GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020, and contrasts with the Maze results, where the 5-gram model had the worst fit and did not provide additional predictive value over the other models. The fact that the neural language models, which capture deeper features of linguistic structure  [@huSystematicAssessmentSyntactic2020,@wilcox-etal:2022-using-computational-models], dominate the more superficial $n$-gram models for Maze but not SPR suggests that Maze RTs reflect richer language structure-related processes during real-time comprehension than do SPR RTs.

@@

Fourth, we compared the spillover profiles between Maze and SPR. For Maze, we found large effects of the current words surprisal and length, which dwarfed any spillover effects from previous word predictors. In contrast, for SPR, we found effects of roughly equal sizes from the current and previous words.[^Furthermore, the typical spillover profile for SPR data may be worse than suggested by the Natural Stories corpus SPR data: for example, @smithEffectWordPredictability2013, found that most of a word's surprisal effect showed up only one to two words downstream.] Overall, Maze is a slower task than SPR, but it also has much larger effects of length and surprisal, perhaps due to requiring more focus, and thus generating less noisy data.

Lastly, we examined how different language models fare at predicting human RT data. We found that overall, the models were more predictive of the A-maze data than SPR data; however, the hierarchy of the model's predictive performance also differed between the A-maze and SPR datasets. This difference suggests that how well a language model predicts human RTs may depend on task. Maze RTs were by far best predicted by language models that capture rich features of natural language grammatical structure, unlike SPR RTs which were nearly equally well predicted by more superficial $n$-gram models. These findings further add to the evidence that the Maze task is favorable for RT-based investigations of underlying linguistic processing in the human mind. More broadly, further comparisons between different processing methods on the same materials could be useful for a deeper understanding of how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

Overall, A-maze has excellent good localization, although some  models showed small but statistically reliable effects of the past word. On the whole, our results support the idea that Maze forces language processing to be close to word-by-word, and thus the Maze task can be used under the assumption that the RT of a word primarily reflects its own properties and not those of earlier words.

@@

Compared to traditional Maze, error-correction Maze reduces perverse (from the experimenter's point of view) incentives to complete the task as quickly as possible. However, even with error-correction Maze, clicking randomly is still likely faster than doing the task. In discussing this work, we received the suggestion that one way to further disincentivize random clicking would be to add a pause when a participant makes a mistake, forcing them to wait some short period of time (ex 500ms) before correcting their mistake. This delay would make randomly hitting buttons slower than doing the task as intended.

Error-correction Maze records RTs for words after a participant makes a mistake in the sentence. In our analyses, we excluded these post-error data, but we believe it is an open question whether data from after a participant makes a mistake is usable. That is, does it show the same profile as RTs from pre-error words, or are there traces from recovering from the mistake? If there are, how long do these effects take to fade? Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is hard to assess; if it can be used, it would make the Maze task more data efficient.

The Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest. Its forced incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.  We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods, and we suggest that the error-correction mode be the default choice for presenting Maze materials.
